\documentclass[english,12pt]{revtex4}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{nicefrac}

\begin{document}
\title{{\Large Coulomb correction to the thermodinamic parameters of dense plasmas}}
\maketitle

{\bf Coulomb interactions} we account within the Madelung approximation. The extra term in the free energy which accounts for the
electrostatic energy of each of the ions in the charge state $i$ coupled with i electrons, the latter being uniformly
distributed over the "iono-sphere"
(see Eq.(3.50) in \cite{drake}):
\begin{equation}\label{fterm}
F_M=-\frac{9}{10} \frac{q_e^2}{r_{iono}} \sum_{i=0}^{i_{max}} i^2 N_i,\qquad
r_{iono} = \left( \frac{4 \pi}{3} n_a \right)^{-\frac13}.
\end{equation}
Accordingly, in the requirement of the ionization equilibrium, $\partial F/\partial N_i - \partial F/\partial N_{i+1} - \partial F/\partial N_e = 0$ (with respect to the reaction $(i)\leftrightarrow(i+1)+e$),
the term $\partial F_{M}/\partial N_i -\partial F_{M}/\partial N_{i+1}$ will 
give the contribution of $\frac{9}{10} \frac{(2i+1)q_e^2}{r_{iono}}$ into the left hand side:
\begin{equation}
-T \log \left[ \frac{g_i}    {N_i}     e^{\sum_{j=0}^{i-1} I_j/T} \right] - \frac{9}{10} \frac{i^2     q_e^2}{r_{iono}}
+T \log \left[ \frac{g_{i+1}}{N_{i+1}} e^{\sum_{j=0}^i     I_j/T} \right] + \frac{9}{10} \frac{(i+1)^2 q_e^2}{r_{iono}}
-\mu_e = 0.
\end{equation}
The solution of the ionization equilibrium, hence, reads:
\begin{equation}
\frac{N_{i+1}}{g_{i+1}} = \frac{N_i}{g_i} \exp\left(-\frac1T \left(I_i - \frac{9}{10} \frac{(2i+1) q_e^2}{r_{iono}} + \mu_e \right)\right),
\end{equation}
or, applying this recursively and reducing the sum $\sum_{j=0}^{i-1} (2j+1) = i^2$,
\begin{equation}\label{pfM}
\frac{N_i}{g_i}=\frac{N_0}{g_0}(g_e)^i \exp \left( \frac{9}{10} \frac{i^2 q_e^2}{T r_{iono}} -\frac{\sum_{j=0}^{i-1}I_j}T \right) .
\end{equation}
This may be interpreted as the ionization potential lowering which results from the Coulomb interaction:
each of the potentials, $I_i$, is reduced by $\frac{9}{10} \frac{(2i+1)q_e^2}{r_{iono}}$.
Energy of the ion of the charge state $i$ is $E_i^* = \sum_{j=0}^{i-1}I_j - \frac{9}{10} \frac{i^2 q_e^2}{r_{iono}}$.
This effect shifts the ionization equilibrium towards higher ionization degrees, for a given temperature and atomic density.

The common multiplier, $\frac{N_0}{g_0}$, in each of Eqs.(\ref{pfM}) may be also represented as $\frac{N_a}{S}$.
From the normalization condition, $\sum N_i = N_a$, we find that $S$ is a statistical sum:
\begin{equation}
S=\sum_{i=0}^{i_{max}} g_i (g_e)^i \exp\left(-\frac{E_i^*}T\right),
\end{equation}
so that:
\begin{equation}\label{ni}
p_i = \frac{N_i}{N_a} = \frac{1}S g_i (g_e)^i \exp \left( -\frac{E_i^*}T \right).
\end{equation}

The full Helmholtz free energy now includes the contribution of the electrostatic field energy as in Eq.(\ref{fterm}):
\begin{equation}\label{fe1}
F=-T
\sum_{i=0}^{i_{max}}{
N_i\log\left[g_i
  \frac{eV}{N_i}\left(\frac{MT}{2\pi \hbar^2}\right)^{3/2}\exp \left(-\sum_{j=0}^{i-1}\frac{I_j}T \right)\right]}+F_e
  -\frac{9}{10} \sum_{i=0}^{i_{max}} N_i \frac{i^2 q_e^2}{r_{iono}}.
\end{equation}  

With the ion partition functions as in Eq.(\ref{ni}) one can rewrite Eq.(\ref{fe1}) in the following form:
\begin{equation}\label{ffullm}
F = -TN_a\log\left[\frac{eV}{N_a}\left(\frac{MT}{2\pi \hbar^2}\right)^{3/2}\right]-TN_a\log S + \Omega_e, 
\end{equation}
where, again, $\Omega_e = F_e - \mu_e N_a \sum i p_i = F_e - \mu_e N_a \langle i \rangle $.

{\bf Differentials along the curve of the ionization equilibrium} obey the equation as follows:
\begin{equation}
A_{g_e} \frac{dg_e}{g_e} = A_V \frac{dV}{V} + A_T \frac{dT}{T}
\end{equation}
\begin{equation}
A_{g_e} = \langle \delta^2 i \rangle + ZR^-(g_e), \qquad
A_T     = \frac32 Z - \frac{\langle \delta i \delta E^*_i \rangle}{T}, \qquad
A_V     = Z + \frac{3}{10} \frac{q_e^2}{T r_{iono}} \langle \delta(i^2) \delta i \rangle
\end{equation}
Again, we express the result in terms of covariances,
$\langle \delta a \delta b \rangle = \langle (a - \langle a \rangle) (b - \langle b \rangle) \rangle$,
and mean values which are now being calculated
using the modified partition functions.
Differentiation of mean values which is necessary for derivation of the above equation on
differentials is not a complicated problem with the following formula:
\begin{equation}
d \langle f_i \rangle = \left\langle \delta f_i \delta\left( \frac{dp_i}{p_i} \right) \right\rangle,
\end{equation}
where $f_i$ is a function of the only argument $i$, for example, $iE_i$ or $i^2+i$.

{\bf Plasma thermodynamics and Equation-Of-State.} 
While differentiating Eq.(\ref{ffullm}) with respect to $T$ and $V$, again, we see that the derivatives
by $g_e$ from the second and third terms cancel 
each other: $g_e(\partial \log S/\partial g_e)=\langle i\rangle=Z$,
that is evident from $d \log S = \langle d (\log p_i) \rangle$,
and $-g_eg_{e1}{\rm Fe}^\prime_{3/2}(g_e)=g_{e1}{\rm Fe}_{1/2}=Z$. That is why for the internal energy density, 
${\cal E}$, and for the pressure, $P$, we find:
\begin{equation}
{\cal E} = -\frac{T^2}V\left(\frac{\partial}{\partial T}\left(\frac F T\right)\right)=
{\cal E}_i+{\cal E}_e,\qquad{\cal E}_i=\frac32Tn_a,\qquad
{\cal E}_e = n_a\left[ \frac32 T Z R^+ + \langle E^*_i\rangle \right],
\end{equation}
\begin{equation}
P = -\frac{\partial F}{\partial V}=P_i+P_e,\quad
P_i = n_aT,\quad
P_e = n_aT ( 1 + ZR^+ - L \langle i^2 \rangle ),\quad
L = \frac{3}{10} \frac{q_e^2}{Tr_{iono}}.
\end{equation}
In the above equations we add the Madelung corrections into the energy of the electron gas, ${\cal E}_e$, and
into the pressure of electrons, $P_e$, because those corrections are controlled by the electron temperature.

However, while calculating the second order thermodynamic derivatives,
such as specific heat, the derivatives of $g_e$ essentially sophisticate the 
calculations. 

In such a way one can find the specific heat in isochoric process, per unit of volume:
\begin{equation}
C_{Ve}=\frac{\partial {\cal E}_e}{\partial T}=n_a\left[\frac{\langle\delta^2E^*_i\rangle}{T^2}+\frac{15}4ZR^+
-\frac{\left(\frac32Z-\frac{\langle\delta E^*_i\delta i\rangle}T\right)^2}{\langle\delta^2i\rangle+ZR^-}\right],
\end{equation}
the temperature derivative of pressure:
\begin{equation}
\frac {\partial P_e}{\partial T}=
n_a\left[
	\frac52 Z R^+ -
	(Z+L \langle \delta(i^2) \delta i \rangle)
		\frac{\frac32Z-\frac{\langle\delta E^*_i\delta i\rangle}T}{\langle\delta^2i\rangle+ZR^-} -
	L \frac{\langle \delta E^*_i \delta(i^2) \rangle}T
\right],
\end{equation}
as well as the isothermal compressibility:
\begin{equation}
V\frac{\partial P_e}{\partial V}=
n_a T \left[ -\frac{(Z + L \langle \delta(i^2) \delta i \rangle)^2}{\langle \delta^2 i \rangle + ZR^-} +
L \left( \frac43 \langle i^2 \rangle + L \langle \delta^2(i^2) \rangle \right) \right].
\end{equation}
For simplicity in the above equations the contributions due to ion translational motions are omitted, which are:
\begin{equation}
C_{Vi}=\frac32n_a, \qquad
\frac{\partial P_i}{\partial T}=n_a, \qquad
V\frac{\partial P_i}{\partial V}=-n_aT.
\end{equation}

Following is the table showing the values of $Z$ calculated for Xenon
at various electron temperatures
(given in electron-volts -- the value of $k_{B}T_{e}$, where $k_{B}$
is in eV/K) and heavy particle concentrations (given in number of
particles per $cm^{3}$). "no" columns contain the same values
calculated without Coulomb interation taken into account.

\begin{center}
\input{ionization_table7.tex}
\par\end{center}


%\include{FermiFunction}
%\include{CoulombInteraction}


%{\bf Conclusion.} Again, mention that in the partially ionized plasma the exchamge interaction
%effect on the plasma pressure is not exhausted by the multiplier  $R^+\ge 1$, there is also the reduction in $Z$ proportional to negative 
%$\delta_{\rm Fe}$. This thermodynamical property of the partial ionized gas should be also taken into account while 
%generalizing the result for the case of small but finite Coulomb interactions, which may be  done by developing the
%chemical potential of the electron gas into a series of the average Coulomb potential of electrons.   



\begin{thebibliography}{99}
\bibitem{drake}
R.P.Drake. High Energy Density Physics. Springer, Berlin-Heidelberg-NY, 2006.
\bibitem{AA2}
J.Stein, D.Shalitin, and Akiva Ron. Average-atom models of line broadening in hot dense plasmas.
Physical Review A, Volume 31, Number 1, 1985.
\bibitem{CollisionRadiative}
D.Wunderlich, S.Dietrich, U.Fantz.
Application of a collisional radiative model to atomic hydrogen for diagnosic purposes.
JQS\&RT 110 (2009) 62-71.
\bibitem{CollRad}
J.Yuan, G.A.Moses. YAC: A code using the detailed term accounting model for all-Z elements.
JQS\&RT 99 (2006) 697-711.
\bibitem{EOSTA}
A.Bar-Shalom, J.Oreg, M.Klapisch.
EOSTA -- an improved EOS quantum mechanical model in the STA opacity code.
JQS\&RT 99 (2006) 35-54.
\bibitem{InfernoLike}
B.Wilson, V.Sonnad, P.Sterne, W.Isaacs. PURGATORIO -- a new implementation of the INFERNO algorithm.
JQS\&RT 99 (2006) 658-679.
\bibitem{LineBroadening}
E.Stambulchik, Y.Maron.
A study of ion-dynamics and correlation effects for spectral line broadening in plasma: K-shell lines.
JQS\&RT 99 (2006) 730-749.
\bibitem{LineByZLine}
M.Lino da Silva.
An adaptive line-by-line -- statistical model for fast and accurate spectral simulations in low-pressure plasmas.
JQS\&RT 108 (2007) 106-125.
\bibitem{NonLTEXeSpectrum}
J.Bauche, C.Bauche-Arnoult, O.Peyrusse, A.Bachelier, K.B.Fournier, C.Chenais-Popovics, J.-C.Gauthier
Analysis of a non-LTE xenon spectrum by means of the model of superconfiguration temperatures.
JQS\&RT 81 (2003) 47-55.
\bibitem{Rodriguez0}
R.Rodriguez, J.M.Gil, J.G.Rubiano, R.Florido, P.Martel, E.Minguez.
Relativistic quantum mechanic calculation of photoionization cross-section of hydrogenic and
non-hydrogenic states using analytical potentials.
JQS\&RT 91 (2005) 393-413.
\bibitem{RodriguezPhotoIon}
R.Rodriguez, J.M.Gil, R.Florido.
Photoionization cross section of non-hydrogenic levels for weakly coupled plasmas.
JQS\&RT 108 (2007) 239-255.
\bibitem{STA}
A.Bar-Shalom, J.Oreg, W.H.Goldstein, D.Shvarts, A.Zigler.
Super-transition-arrays: A model for the spectral analysis of hot, dense plasma.
Physical Review A, Volume 40, Number 6, 1989.
\bibitem{StewartPyatt}
J.C.Stewart, K.D.Jr.Pyatt.
Lowering of ionization potentials in plasmas.
Astrophysical Journal, Volume 144, Number 3, p.1203, 1966.
\bibitem{SuperConfiguration}
Jean-Christophe Pain, Thomas Blenski.
Self-consistent approach for the thermodynamics of ions in dense plasmas
in the super configuration approximation.
JQS\&RT 81 (2003) 355-369.
\bibitem{Voight}
Mofreh R.Zaghloul.
Comment on "a fast method of modeling spectral line".
JQS\&RT 109 (2008) 2895-2897.
\bibitem{WireArray}
M.F.Yilmaz, A.S.Safronova, V.L.Kantsyrev, A.A.Esaulov, K.M.Williamson, G.C.Osborne, I.Sherestha, N.D.Ouart.
Spectroscopic features of implisions of Mo single- and double-planar wire arrays produced on the 1 MA Z-pinch generator.
JQS\&RT 109 (2008) 2877-2890.
\bibitem{Xe11}
H.-P.Garnir, S.Enzonga Yoca, P.Quinet, E.Biemont.
Lifetime and transition probability determination in Xe IX.
JQS\&RT 110 (2009) 284-292.
\end{thebibliography}

\end{document}







{\bf In the limit of strong plasma non-ideality,} specifically, if the 'ionosphere radius', $a=(3/4\pi n_a)^{1/3}$, is not small as compared with the Debye radius: 
$a\ge R_D$, the  Debye-Huekel theory is not applicable. While treating this limiting case, however, it should be mentioned, that if the condition $a\gg R_D$, or, the same,
$a\ll Z^2q_e^2/T$ is achieved at the cost of high atomic concentration, then the pressure ionization becomes dominant. The latter effect is significant, if the ionosphere 
radius is less than the radius of the dominant ion as in the Thomas-Fermi model: $a\le R_{TF}$. Under this condition the approach of the present would be not directly 
applicable, because the ion energy levels, $E_i$ in this case are strongly affected by the compression, resulting ultimately in the total disappearance of the bound 
states for electron in the limit of highest pressures.  To neglect this effect we have to assume that the density is limited from above:
\begin{equation}
a\gg R_{TF}(Z,i_{\max}).
\end{equation}  
Under this condition, which is essentially the condition for the thermal ionization dominating the pressure ionization, the plasma strong non-ideality may occur only at low 
temperature. In this case we can use the Madelung estimate for the electrostatic energy, $-9i^2 q_e^2/10a$ - see \cite{PD} for more detail. This estimate seems to be the
asymptotic exppression for high $i\gg1$, particularly, for $i=1$ the positive contribution fron the electron-electron interaction estimated as $\sim +i^2$ 
should in fact turn to zero ($=i(i-1)$?). Therefore, while considering the transition from the Debye-Huekel electrostatic interaction energy per atom,  
$-(i^2+i)q_e^2/2R_D$, to the Madelung one we evaluate 
the latter for simlicity as as $-9(i^2+i)q_e^2/10a$. To achieve this, the only thing we pratically need is to change $R_D\rightarrow \max(R_D,5a/9)$ in 
Eqs.(\ref{FDHApprox},\ref{ZDHApprox},\ref{EDHApprox},\ref{PDHApprox}).


