\begin{thebibliography}{{\em Moen and Brekke}(1993}

\bibitem[{\em Goodman}(1995)]{goodman95}
Goodman, M.~L.,
\newblock A three-dimensional, iterative mapping procedure for the
  implementation of an ionosphere-magnetosphere anisotropic {Ohm's} law
  boundary condition in global magnetohydrodynamic simulations,
\newblock {\em Ann. Geophysic.}, {\em 13}, 843, 1995.

\bibitem[{\em Manchester et al.}(2004)]{manchester_cme} Manchester IV,
W. B., T. I. Gombosi, I.I Roussev, D.L. De Zeeuw, I.V. Sokolov,
G. Toth, M. Opher, K.G. Powell \newblock Three-Dimensional MHD
Simulation of a Flux Rope Driven CME, \newblock {\em
J. Geophys. Res.}, {\em 109}, A01102, doi: 10.1029/2002JA009672, 2004.

\bibitem[{\em Moen and Brekke}(1993)]{moen93}
Moen, J., and A.~Brekke,
\newblock The solar flux influence os quiet-time conductances in the auroral
  ionosphere,
\newblock {\em Geophys. Res. Lett.}, {\em 20}, 971, 1993.

\bibitem[{\em Powell et~al.}(1999)]{powell99}
Powell, K.G., P.L. Roe, T.J. Linde, T.I. Gombosi, and D.L.~De Zeeuw,
\newblock A solution-adaptive upwind scheme for ideal magnetohydrodynamics,
\newblock {\em J. Comp. Phys.}, {\em 154}, 284, 1999.

\bibitem[{\em Ridley et~al.}(2001)]{ridley_aec}
Ridley, A.J., D.L.~De Zeeuw, T.I. Gombosi, and K.G. Powell,
\newblock Using steady-state mhd results to predict the global state of the
  magnetosphere-ionosphere system,
\newblock {\em J. Geophys. Res.}, {\em 106}, 30,067, 2001.

\bibitem[{\em Ridley et~al.}(2002a)]{ridley_coupled}
Ridley, A.J., T.I. Gombosi, D.L.~De Zeeuw, C.R. Clauer, and A.D. Richmond,
\newblock Ionospheric control of the magnetospheric configuration: Neutral
  winds,
\newblock {\em J. Geophys. Res.}, page To be submitted, 2002a.

\bibitem[{\em Ridley et~al.}(2002b)]{ridley_metrics}
Ridley, A.J., K.C. Hansen, G.~T\'oth, D.L.~De Zeeuw, T.I. Gombosi, and K.G.
  Powell,
\newblock University of michigan mhd results of the ggcm metrics challenge,
\newblock {\em J. Geophys. Res.}, page In Press, 2002b.

\bibitem[{\em Robinson et~al.}(1987)]{robinson87}
Robinson, R.M., R.R. Vondrak, K.~Miller, T.~Dabbs, and D.A. Hardy,
\newblock On calculating ionospheric conductances from the flux and energy of
  precipitating electrons,
\newblock {\em J. Geophys. Res.}, {\em 92}, 2565, 1987.

\end{thebibliography}
