%  Copyright (C) 2002 Regents of the University of Michigan, portions used with permission 
%  For more information, see http://csem.engin.umich.edu/tools/swmf
\section{Configuration of PWOM}

Configuration refers to several different ways of controlling how the 
PWOM is compiled and run.  The most obvious is the setting of
compiler flags specific to the machine and version of FORTRAN
compiler.  The other methods refer to ways in which planet the PWOM 
is configured to solve.

\subsection{Setting compiler flags in Makefile.conf and with Config.pl}

The compiler flags can be modified by editing
\begin{verbatim}
  Makefile.conf
\end{verbatim}
This makefile is created during installation, and it contains the
platform and compiler specific part of the makefile system.
The most usual changes can be easily done with the Config.pl script.
The precision of real numbers can be set with
\begin{verbatim}
  Config.pl -single
\end{verbatim}
or 
\begin{verbatim}
  Config.pl -double
\end{verbatim}
If the precision is modified, the script will execute 'make clean',
so that all the files are compiled with the new real precision.

The debugging flags can be switched on and off with
\begin{verbatim}
  Config.pl -debug
\end{verbatim}
and
\begin{verbatim}
  Config.pl -nodebug
\end{verbatim}
respectively. The maximum optimization level can be set to -O2 with
\begin{verbatim}
  Config.pl -O2
\end{verbatim}
The minimum level is 0, the maximum is 4 (Note that not all compilers support
level 4 optimization).

\subsection{Planet Configuration}

The PWOM is set to have a very general core of code, and planet specific 
files that define the chemistry, neutral and ion species and other 
planet parameters. The planet specific codes are stored in srcEarth or 
srcSaturn. When the PWOM is confifugured for a specific planet, the planet 
specific codes are copied into the src directory. To configure the PWOM for 
earth simply type
\begin{verbatim}
  Config.pl -Earth
\end{verbatim}
to configure for Saturn
\begin{verbatim}
  Config.pl -Saturn
\end{verbatim}


